### 2021-05-05 analysis/statistical_analyses.Rmd
revised entire paper

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)
library(here)
library(ggthemes)
library(lme4)
library(brms)
library(xtable)
```

```{r data, include=TRUE}
dpath_exp1 <- here("data", "responsesOnly_E1_100def.csv")
dpath_exp2 <- here("data", "responsesOnly_E2_100def.csv")

mytheme <-  theme_classic(base_size = 10)  + 
            theme(strip.background = element_blank())

theme_set(mytheme)
```

```{r loadds, include=TRUE}

# Experiment 1 data frame 
d1 <- read_csv(dpath_exp1) #%>%
d2 <- read_csv(dpath_exp2) #%>%

```

### Plot data

```{r exp1plot, include=TRUE, fig.height=4, fig.width=6}

d1plot <- d1 %>% 
  filter(stimulustype != "chain") %>% 
  group_by(state_or_event, length, response) %>% 
  summarize(count = n()) %>% 
  mutate(response = factor(response, levels = c("RC_IC", "RC", "RD_ID", "RD", "IC", "other")))

plot1 <- d1plot %>% 
  ggplot(aes(x = response, y=count)) +
  geom_bar(stat="identity") +
  facet_grid(state_or_event~length)
plot1

```

### Stats for Expt 1

```{r exp1stats, eval=TRUE, fig.height=4, fig.width=6}

d1_stats1 <- d1 %>% 
  mutate(ri = factor(case_when(response=="RC_IC" ~  NA_character_,
                        response=="RC" ~     "root",
                        response=="RD_ID" ~  NA_character_,
                        response=="RD" ~     "root",
                        response=="IC" ~     "immediate",
                        response=="ID" ~     "immediate",
                        TRUE ~ NA_character_ 
                        ), levels = c("immediate", "root")),
         cd = factor(case_when(response=="RC_IC" ~  "continuous",
                        response=="RC" ~     "continuous",
                        response=="RD_ID" ~  "delayed",
                        response=="RD" ~     "delayed",
                        response=="IC" ~     "continuous",
                        response=="ID" ~     "delayed",
                        response=="otherC" ~     "continuous",
                        response=="otherD" ~     "delayed",
                        TRUE ~ NA_character_ 
                        ), levels = c("delayed", "continuous")),
         length = factor(length, levels = c("short", "medium", "long")),
         state_or_event = factor(state_or_event, levels = c("event", "state")),
         position = factor(case_when(is.na(position) ~  "middle",
                                     TRUE ~ position
                        ), levels = c("top", "bottom", "middle")),
         delay = factor(case_when(is.na(delay) ~  "none",
                                     TRUE ~ as.character(delay) 
                        ), levels = c("none", "2", "4", "6"))
        )

d1_stats1ri <- d1_stats1 %>% 
  drop_na(ri) %>% 
  drop_na(state_or_event) # drop chain

d1_stats1cd <- d1_stats1 %>% 
  drop_na(cd) %>% 
  drop_na(state_or_event)  # drop chain

# use sum coding 
#options(contrasts=c("contr.sum", "contr.poly"))
options(contrasts=c("contr.treatment", "contr.poly"))

m0 <- glmer(ri ~ state_or_event + length + (1|participantnum), data = d1_stats1, 
            family = binomial(link = "logit"))

nchain <- 1 # for testing
niter <- 500 # for testing

nchain <- 8 # for real run
niter <- 4000 # for real run

fit1_ri <- brm( ri ~ state_or_event + length + (1|participantnum), family=bernoulli(), data = d1_stats1ri, chains = nchain, iter = niter, control = list(adapt_delta = 0.99) )

fit1_cd <- brm( cd ~ state_or_event + length + (1|participantnum), family=bernoulli(), data = d1_stats1cd,  chains = nchain, iter = niter )

cd_tab <- fixef(fit1_cd) %>% as_tibble(rownames="NA")

mcdtableout<- xtable(cd_tab, caption = "Coefficients for cd regression, Expt 1" )
print(mcdtableout, file = here("analysis", "output", paste0("coeftable_cd_exp1.tex")))

ri_tab <- fixef(fit1_ri) %>% as_tibble(rownames="NA")

mritableout<- xtable(ri_tab, caption = "Coefficients for ri regression, Expt 1" )
print(mritableout, file = here("analysis", "output", paste0("coeftable_ri_exp1.tex")), include.rownames = FALSE)

```

### Fig 2
```{r exp2plot, include=TRUE, fig.height=4, fig.width=6}

d2plot <- d2 %>% 
  filter(stimulustype != "chain") %>% 
  group_by(order, response) %>% 
  summarize(count = n()) %>% 
  mutate(response = factor(response, levels = c("RC", "RD", "IC", "ID", "other")))

plot2 <- d2plot %>% 
  ggplot(aes(x = response, y=count)) +
  geom_bar(stat="identity") +
  facet_wrap(~order)
plot2

```

### Stats for Experiment 2

```{r exp2stats, eval=TRUE, fig.height=4, fig.width=6}

d2_stats1 <- d2 %>% 
  mutate(ri = factor(case_when(response=="RC" ~     "root",
                               response=="RD" ~     "root",
                               response=="IC" ~     "immediate",
                               response=="ID" ~     "immediate",
                               TRUE ~ NA_character_ 
                               ), levels = c("immediate", "root")),
         cd = factor(case_when(response=="RC" ~     "continuous",
                               response=="RD" ~     "delayed",
                               response=="IC" ~     "continuous",
                               response=="ID" ~     "delayed",
                               response=="otherC" ~     "continuous",
                               response=="otherD" ~     "delayed",
                               TRUE ~ NA_character_ 
                               ), levels = c("delayed", "continuous")),
         order = factor(order, levels = c("RC_RD_ID", "RD_RC_ID", "RD_ID_RC"))
        )

d2_stats1ri <- d2_stats1 %>% 
  drop_na(ri)

d2_stats1cd <- d2_stats1 %>% 
  drop_na(cd)

# use sum coding 
#options(contrasts=c("contr.sum", "contr.poly"))
options(contrasts=c("contr.treatment", "contr.poly"))

m0 <- glmer(ri ~ order + (1|participantnum), data = d2_stats1, 
            family = binomial(link = "logit"))

nchain <- 1 # for testing
niter <- 500 # for testing

nchain <- 4 # for real run
niter <- 2000 # for real run

fit2_ri <- brm( ri ~ order + (1|participantnum), family=bernoulli(), data = d2_stats1ri, chains = nchain, iter = niter )

fit2_cd <- brm( cd ~ order + (1|participantnum), family=bernoulli(), data = d2_stats1cd,  chains = nchain, iter = niter )

cd_tab <- fixef(fit2_cd) %>% as_tibble(rownames="NA")

mcdtableout<- xtable(cd_tab, caption = "Coefficients for cd regression, Expt 2" )
print(mcdtableout, file = here("analysis", "output", paste0("coeftable_cd_exp2.tex")))

ri_tab <- fixef(fit2_ri) %>% as_tibble(rownames="NA")

mritableout<- xtable(ri_tab, caption = "Coefficients for ri regression, Expt 2" )
print(mritableout, file = here("analysis", "output", paste0("coeftable_ri_exp2.tex")), include.rownames = FALSE)

```